

Consider a quick example matrix

```{r}
M <- matrix(rnorm(96), nrow=8) #  12 by 8 

```

The neighborhood you describe is what we call the [Moore Neighborhood]()  We can generalize it to a Manhattan radius `r` giving the number of cells to go out.  Implementing from the definition, first let us grab a random cell:



```{r}
i <- sample(1:dim(M)[1], 1)  # 1 random number between 1 and 12
j <- sample(1:dim(M)[2], 1)  # another random number between 1 and 8
```

And then grab all the points in the Moore neighborhood 

```{r}
r <- 1
neighborhood <- 
cbind(c(i-r, i,   i+r, i-r, i,  i+r,  i-r, i,   i+r),
      c(j-r, j-r, j-r, j,   j,  j,    j+r, j+r, j+r))
```

We can then access the cells from M in our neighborhood.  We tell R to display these points as a matrix 

```{r}
matrix(M[neighborhood], nrow = 3)
```

 In general there are concerns about what to do when `i,j` sits at the boundary.  One option is to wrap that around the edge (effectively treating things like a torus).  We use the modulus function, `%%`.  We add and subtract `1` because this trick works by indexing vectors starting with 0, not 1:

```{r}
wrapped_neighborhood <- neighborhood
wrapped_neighborhood[,1] <- (neighborhood[,1] - 1) %% dim(M)[1] + 1
wrapped_neighborhood[,2] <- (neighborhood[,2] - 1) %% dim(M)[2] + 1
```


This can be extended to $r > 1$. 

```{r}
r_max <- 3
rows <- numeric(0)
cols <- numeric(0)

for(r in 1:r_max){
 rows <- c(rows, c(i-r, i,   i+r, i-r, i,  i+r,  i-r, i,   i+r))
 cols <- c(cols, c(j-r, j-r, j-r, j,   j,  j,    j+r, j+r, j+r))
} 

neighborhood <- cbind(rows, cols)
```

Once again we have to deal with boundary cases, e.g. wrapping as before:  

```{r}
wrapped_neighborhood <- neighborhood
wrapped_neighborhood[,1] <- (neighborhood[,1] - 1) %% dim(M)[1] + 1
wrapped_neighborhood[,2] <- (neighborhood[,2] - 1) %% dim(M)[2] + 1
```



What if we wanted to wrap just the neighborhood if we hit the boundary?

```{r}
r_max <- 3
rows <- numeric(0)
cols <- numeric(0)

wrap <- function(i, r, n){
 if(i-r < 0)
   out <- i+r+1
 if(i+r > n)
   out <- i-r-1
 out
}

for(r in 1:r_max){
 rows <- c(rows, c(i-r, i,   i+r, i-r, i,  i+r,  i-r, i,   i+r))
 cols <- c(cols, c(j-r, j-r, j-r, j,   j,  j,    j+r, j+r, j+r))
 rows <- wrap(rows, i, r, dim(M)[1])
 cols <- wrap(cols, j, r, dim(M)[2])
} 

neighborhood <- cbind(rows, cols)
```


Note that this does not handle the case of a neighborhood that is larger than the matrix in some dimension; e.g. we don't guarentee $i+r+1<n$, etc.  

